Project-I by Group CITYNAME

### 2015-10-27 report/paper.tex
First part of the report written and compiled.

\author{
Corentin Tallec\\
Roman Shirochenko\\
EPFL \\
\texttt{corentin.tallec@epfl.ch\\
roman.shirochenko@epfl.ch} \\
}

@@

\subsection{Data visualization and cleaning}
We performed basic exploratory data analysis on our data. Figure \ref{fig:boxplotX} shows distributions of all 10 input variables. As expected the data is not centered, and therefore we normalize the data. Figure \ref{fig:histY} shows the histogram of the output variable. We see that there are two data points which have rather high value of the output variables compared to rest of the data. Since these are only two data points, we remove them from the data (example number 362 and 402).

We also investigated the correlation between output and input variables (plot not shown, since it is not important for our analysis). We found that most of the variables are correlated with the output, confirming the need for a full regression analysis. The third variable did not seem significantly correlated, however this preliminary observation does not suggest removal of this variable from the dataset.

We use a dummy encoding for the categorical variable. The binary variable does not require any dummy encoding. This gives us a total of 14 input variables.

The input matrix $\mathbf{X}$ is rank-deficient with a rank of 6, instead of 10. This implies that we must ``lift" the eigenvalues, using methods such ridge regression.

\begin{figure}[!t]
\center
\subfigure[Boxplot of real-valued $\mathbf{X}$. Data is not centered and therefore we normalize it.]{\includegraphics[width=2.5in]{figures/boxplotX.pdf} \label{fig:boxplotX}}
\hfill
\subfigure[Histogram of $\mathbf{y}$. We can clearly see two outliers.]{\includegraphics[width=2.5in]{figures/histY.pdf} \label{fig:histY}}
\caption{}
\end{figure}

\subsection{Ridge regression}
We applied least-squares and ridge regression to this dataset. Since the matrix is ill-conditioned, least-squares is not suitable. Therefore, we report results obtained with ridge regression alone. Note that the improvements using ridge regression were modest and not much lower than that of linear regression, however we do not expect least-squares to work well when there is a lot of testing data available.

Figure \ref{fig:ridgeCurve} shows the results obtained with ridge regression when we use $50\%$ of the data as test data and rest as training data. We varied the value of $\lambda$ from $10^{-4}$ to $10^3$, choosing total 500 points in between. We can see that there is a small improvement obtained for some values of $\lambda$.

We did experiments to plot a learning curve for this data (see Andrew Ng's notes about the learning curve). We held out 20\% of data as test data and rest as training data. We chose to slowly increase the proportion of data used for training. For each proportion of the training data, we repeated the experiment 30 times to compute the distribution of error. We fit ridge regression to each sampled training set and test it on the same $20\%$ test data. We varied the value of $\lambda$ from $10^{-4}$ to $10^3$, choosing total 500 points in between.

This gives us the learning curve shown in Fig. \ref{fig:learnCurve}. The blue curve shows the train error while red curve shows the test error. We can see that both training and test error converge, with the variance of estimates decreasing as we increase the training data size. There is also a very small gap between the train and test errors, showing that the linear model is a reasonable choice. The small gap exists perhaps because we have only limited test data.

\subsection{Feature transformations}
We tried several feature transformations. We found that we get a small improvement in performance when we take $\sqrt{|X_{ni}|}$ for all entries of $\mathbf{X}$. We did not check (due to lack of time) whether it matters if we apply this to one variable or all. We performed experiments similar to the last section (although one should really do cross-validation). Values of lambda were kept same as the last section.

We compare three methods. First is a baseline where we do not use any input variables i.e. mean value of the output. The second method is the ridge regression described in previous section. The third method is ridge regression with a feature transformation. The first method gave RMSE of around 3 which was way worse than the other two methods.

The RMSE for the last two methods are shown in Fig. \ref{fig:summary}. We see that both test and train error decrease, however it appears that the improvement is very little and may not be significant.

\begin{figure}[!h]
\center
\subfigure[Ridge regression for a 50-50 split.]{\includegraphics[width=2.5in]{figures/ridgeCurve.pdf} \label{fig:ridgeCurve}}
\hfill
\subfigure[Comparison of ridge regression with and without feature transformation. The improvement is very little and might be insignificant.]{\includegraphics[width=2.7in]{figures/summary.pdf} \label{fig:summary}}
\subfigure[Learning curve. Blue is training data and red is test data.]{\includegraphics[width=4in]{figures/learningCurve.pdf} \label{fig:learnCurve}}
\caption{}
\end{figure}

\subsection{Summary}
In this report, we analyzed a regression dataset and found that ridge regression is a reasonable fit. We estimate that the average test error is 1.213 ($\pm$ 0.02). We tried some feature transformation and found that there is a small improvement giving us a test error of around 1.198 ($\pm$ 0.015). This improvement, however, is not significant.

\subsubsection*{Acknowledgments}
We would like to thank Timur for making the dataset for this study, Carlos and Ilija to help conduct the experiments, and all the students who attended the session making it exciting for us. All the code, as well as this report, was written by Emti.

\subsubsection*{References}

>>>>

\author{
Corentin Tallec\\
Roman Shirochenko\\
EPFL \\
\texttt{corentin.tallec@epfl.ch} \\
}

@@

\subsection{Data visualization and cleaning}
The first tasks we performed on the data was to visualize it. The first thing we
did was to obtain a visualization of the data, showed in \ref{fig:boxplotX} and
\ref{fig:histY}. Those two figures led us to normalize both the input and the
output variable (normalizing the output variable allow us to have results that are
easier to interpret, as they are to be compared to values close to 1). No
obvious outliers are noticeable on the $\mathbf{y}$ histogram.

We performed featurewise analysis to notice that most of our real valued
variables have nearly gaussian distributions (or at least unimodal
distributions), which is not the case of our output variable. Figure
\ref{fig:cor1} shows how the output variable evolves as a function of the first
feature. This behaviour is what we notice on most variables, and thus shows no
clear correlations.

Although for variables 49 and 66 we notice a multimodal distribution that is
closer to the behaviour of our output variable, as seen in Figure
\ref{fig:xhist49} and Figure \ref{fig:xhist66}. We thus chose to try and predict
$\mathbf{y}$ using only those two features, transformed through polynomial
transformation.

The categorical variables were encoded using dummy encoding, what led to a total
feature number of 100. As the rank of our dataset is below 100 (it is 70),
simple linear regression is likely to fail, and we use ridge regression to cope
for the rank deficiency.

\begin{figure}[!t]
\center
\subfigure[Boxplot of $\mathbf{X}$.]{\includegraphics[width=2.5in]{figures/boxplot.jpg} \label{fig:boxplotX}}
\subfigure[$\mathbf{y}$ as a function of the first feature of $\mathbf{X}$.
The correlation between the two variables is not
clear.]{\includegraphics[width=2.5in]{figures/cor1.jpg}\label{fig:cor1}}
\hfill
\subfigure[Histogram of
$\mathbf{y}$. The output distribution is bimodal.]{\includegraphics[width=1.5in]{figures/yhist.jpg} \label{fig:histY}}
\hspace*{0.3in}
\subfigure[Histogram of feature 49 of $\mathbf{X}$. We see that the distribution
	is bimodal, as the output.]{\includegraphics[width=1.5in]{figures/xhist49.jpg}
\label{fig:xhist49}}
\hspace*{0.3in}
\subfigure[Histogram of feature 66 of $\mathbf{X}$. Again, the distribution is
	bimodal.]{\includegraphics[width=1.5in]{figures/xhist66.jpg}
\label{fig:xhist66}}
\caption{}
\end{figure}

\subsection{Ridge regression}
As the dimension of the data is inferior to the number of feature, we expect
least-squares to work quite poorly, and thus decided to apply mostly ridge
regression.

For our first attempt at predicting the data, we simply used the initial
features and used kFold with $k=4$ to display the evolution of the RMSE as a
function of the regularizing lambda both for the training and test data. 
The lambda we used ranged from $10^{-4}$ to
$10^3$, with 500 points. The results of the ridge regression are already
satisfactory, with an error of approximately $0.105\pm0.001$ on the training set
and $0.115\pm0.001$ on the test set, that is approximately one tenth of the
variance of the output variable.

We see that for sufficiently low lambda, both the training and test errors do
not vary much (there is indeed only a very slight variation of test RMSE, with a
nearly unnoticeable minimum between $\lambda=10$ and $\lambda=100$).

The results described here are summerized in the Figure \ref{fig:ridgeCurve}.

\subsection{Feature transformations}
As notified in a previous section, among all the feature at our disposal,
features 49 and 66 show distributions that ressemble the ditribution of the
output variable.

This led us to try and regress only on thos two features and polynomial
transformations of them. Using only those two features, we get an RMSE error
that is quite close to the one we notified when all the features are used. When
using polynomial features (up to degree 10), we basically get half the error
that we got using all the features. Those results were obtained using ridge
regression. As in the previous section, we obtained the RMSE as a function of
lambda, calculating both training and test errors using a kFold procedure with
$k=4$. The results are summarized in Figure \ref{fig:polyridgeCurve}, where, as
previously mentionned, we used as features only the two features and their
polynomial transformation with degree ranging from 1 to 10.

This time, the test error varies much more as a function of lambda, and, given
our curve, the best lambda possible seems to be around $10^{-6}$.

Finally to evaluate our performances more accurately, we plotted learning curves
for our model, with 20 features obtained by polynomial transformations and a
lambda of $10^{-6}$. The learning curve is obtained by using $20\%$ of the
training data to perform tests, while learning the model on a progressively
increasing portion of the $80\%$ remaining. We monitor the test and training
error as functions of the proportion of data used to train. The results are
summarized in Figure \ref{fig:learningCurve}.

As expected, both training error and test error decrease as the proportion of
training data increases. The test error obtained seems to exhibit a quit
satisfactory behaviour.

\begin{figure}[!h]
\center
\subfigure[Ridge regression on kFold with
$k=4$.]{\includegraphics[width=2.5in]{figures/traintestRMSE.jpg} \label{fig:ridgeCurve}}
\hfill
\subfigure[Ridge regression on kFold with feature transformation and
	$k=4$.]{\includegraphics[width=2.5in]{figures/traintestRMSEpoly.jpg}
\label{fig:polyridgeCurve}}
\subfigure[Learning curve. Blue is training data and red is test
data.]{\includegraphics[width=4in]{figures/learningcurves.jpg}
\label{fig:learningCurve}}
\caption{}
\end{figure}
